Utilizing Convolution for Obstacle Avoidance during Robotic Manipulator Movement

### 2023-07-27 Articles/ERK/erk.tex
ERK SUBMISSION

This study proposes a method for real-time obstacle avoidance in robotic arm control, leveraging a 3D voxel grid and 3D directional kernels. The approach integrates the benefits of convolutional techniques with the spatial understanding provided by voxel grid representation to efficiently calculate movement directions while evading obstacles. By focusing on kinematics and obstacle grid representation, the method efficiently utilizes the versatility of a 7-degree-of-freedom (DOF) robot arm, employing geometric and differential kinematics for movement calculations. The strategy doesn't require exhaustive training as with neural networks, making it deployable in various environments.

@@

Robotic manipulators, specifically those with a high degree of freedom (DOF), possess a crucial characteristic known as kinematic redundancy~\cite{Nguyen1991}. Kinematic redundancy refers to having more joints (i.e., degrees of freedom - DOF) than required to perform a task, thus allowing the robot to adopt multiple configurations to accomplish the same end-effector position and orientation. This redundancy provides a considerable advantage, making it possible for manipulators to optimize movement according to different criteria, such as energy efficiency, obstacle avoidance, or joint limit avoidance.

@@

When dealing with redundant manipulators, task prioritization is crucial. In our approach, we distinguish between primary and secondary tasks. The primary task of the manipulator is usually defined as reaching a desired EE position and orientation or as following a path made out of such points. In this work we instead use APF (Artificial Potential Field), to guide the EE towards the goal position, while avoiding the obstacles in the path of the EE. Operating in a real-world environment often requires considering secondary tasks, such as real-time obstacle avoidance, to ensure safe and uninterrupted operation of the manipulator.

This paper proposes a novel method for real-time obstacle avoidance of the manipulator arm. We use an obstacle grid combined with convolution techniques to assign avoidance directions to each manipulator segment, ensuring efficient and safe maneuvering within complex environments.

@@

The execution of the secondary task of obstacle avoidance necessitates an understanding of the spatial relationship between the manipulator and surrounding obstacles, usually quantified as a distance or interpreted as a repulsive force. By conceptualizing obstacles as sources of repulsive forces, the manipulator can be guided away from potential collisions, akin to how two magnets of the same polarity repel each other.

Two commonly employed methods for obstacle avoidance include artificial potential field methods (APF)~\cite{Khatib1985} and artificial neural networks. In potential field methods, the robot and obstacles are treated as charged entities. The robot is attracted to the target (goal) while being repelled by obstacles, resulting in a potential field. The robot navigates this field, moving along the path of steepest descent. While this method is intuitive and relatively easy to implement, it can sometimes result in local minima problems where the robot gets stuck in a position that is not the target but cannot find a path due to the repelling obstacles.

@@

The manipulation capabilities of a robot are fundamentally determined by its kinematics, the mathematical description of motion without considering the forces that cause it. In our experiments, we employed a 7 degree-of-freedom (DOF) robot. The seven joints provide a high level of versatility, enabling the robot to move and orient its end-effector in a broad range of positions and directions. To calculate the movements and positions of the robot, we use both geometric and differential kinematics.

Geometric kinematics (also known as forward and inverse kinematics) focuses on the relationship between joint angles and the position and orientation of the end-effector. Forward kinematics involves calculating the position and orientation of the end-effector given the joint angles, while inverse kinematics is the process of determining the joint angles for a desired end-effector position and orientation. In the case or redundant manipulator, there isn't an unique mapping from EE coordinates to the joint coordinates as there exist an infinite number of different joint angles configurations that map to the same EE position.

@@

The null space (eq.~\ref{eq:Null-Space}) of the Jacobian corresponds to those joint motions which produce no movement in the end-effector. The primary task for our robot is the desired end-effector movement. The secondary task of obstacle avoidance in our research takes advantage of this null-space motion~\cite{lajpah_Petri_2012}. Our method applies an avoidance velocity from detected obstacles in the environment. We calculate this velocities at certain points of interest (POI) that are positioned on the manipulator arm segments. We transform avoidance velocities into the joint space velocities using $J^{\dagger}_{poi}$ jacobian calculated in the specific POI. These velocities we than transform into the null-space using matrix $N$. The robot arm is thus able to carry out the primary task while simultaneously avoiding obstacles by appropriately utilizing its redundant DOF (eq.~\ref{eq:manipulator_kinematics}).

@@

We integrated a supplementary secondary task aiming to maintain the robot manipulator's joints as close to the midpoint of their range ( between $q(i)_{max}$ and $q(i)_{min}$ for the i-th joint ) as possible ( eq.~\ref{eq:mid-joints}). $w(i)_m$ weights the importance of each joints position. This auxiliary task, designed to elevate the robot's "elbow", reinforced the robot's upright posture during the execution of the primary task and the obstacle avoidance task.

@@

%\subsection{Avoidance velocities}

%In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations. This methodology allows us to calculate the necessary avoidance velocities for specific points of interest on the manipulator.
%
%The process initiates with kinematic transformations which are used to map the points of interest from the robot's configuration space to the cartesian space and than onto the 3D voxel grid. These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.
%
%Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method. This method is tailored to calculate the constituent parts of the avoidance velocity vector in the Cartesian space, specifically for the x, y, and z directions. For each of these directions, we apply a separate 3D convolutional kernel.

%\todocomment{convolution equation}

%For each direction, we utilize its corresponding kernel and extract a segment of the obstacle grid of matching size, centered at the point of interest. This leads to two equivalent-sized 3D matrices - one from the obstacle grid and the other being the kernel. We then conduct an element-wise multiplication operation between these matrices, summing the resulting values. The result of this operation provides us with the avoidance velocity in each respective direction. Through this approach, we can dynamically adjust the robot's trajectory to effectively avoid obstacles.

@@

In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations. This methodology allows us to calculate the necessary avoidance velocities for specific points of interest located on the manipulator.

The process commences with kinematic transformations, which facilitate the mapping of points of interest from the robot's configuration space to the Cartesian space, followed by their projection onto the 3D voxel grid.

%These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.

@@

Figure~\ref{y axis avoidance values} illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis. Here, the y-coordinates are presented on the x-axis, and the corresponding velocities are plotted on the y-axis. As expected, the y-velocity is the only component exhibiting non-zero values. It effectively propels the robot away from the wall, moving in either a positive or negative direction, depending on the relative position to the wall. Conversely, the x and z velocities remain zero, reflecting the absence of obstacles along these axes. This aligns with our expectations given the presence of the wall obstacle is only along the y direction.

@@

The figure also highlights how the application of the avoidance algorithm prevents the fourth joint from moving closer to the wall, as compared to when the avoidance algorithm is not employed. The position of the joint, upon task completion, is noted to be approximately 20 cm further away from the wall.

@@

%This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and repulsive field velocities for end-effector trajectory generation and robot joint obstacle avoidance. The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.
%
%Notable advantages of our approach include its simplicity and the fact that we work directly in the obstacle space grid, negating the necessity for calculating the spatial primitives of shapes. Moreover, when implemented in parallel, our method exhibits high computational efficiency, making it a viable solution for real-time applications. The experiments conducted effectively showcased that the end-effector could move efficiently towards the target while successfully avoiding obstacles, specifically, a wall in this instance.
%
%Our results affirmed the robustness of the proposed approach, as evidenced by the robot's fourth joint being approximately 20 cm further from the wall upon task completion when the avoidance algorithm was employed.
%
%Despite the promising initial results, our approach does present some limitations. Particularly, the possibility of encountering local minima traps poses a significant challenge. Although these traps could be potentially mitigated by incorporating multiple convolutional kernels of varying sizes, it would result in an approach similar to convolutional neural networks (CNNs), adding to the complexity and computational cost of the algorithm.
%
%Additionally, we acknowledge that our study was conducted under simplified conditions with a single type of obstacle. As such, future work should focus on validating the proposed method in more complex environments, with different obstacle types, shapes, and densities, to further ascertain its adaptability and robustness. It may also be beneficial to explore the integration of other optimization strategies and machine learning techniques to further enhance the overall efficiency and effectiveness of the trajectory planning process.
%
%In conclusion, our work serves as a solid foundation for the development of more advanced and dynamic motion planning algorithms.

@@

Despite its effectiveness, the approach could potentially fall into local minima traps. This limitation could be addressed by incorporating multiple convolutional kernels, though this would increase complexity and resemble convolutional neural networks.

>>>>

This study proposes a method for real-time obstacle avoidance in robotic arm control, leveraging a 3D voxel grid and 3D directional kernels. The approach integrates the benefits of convolutional techniques with the spatial understanding provided by voxel grid representation to efficiently calculate movement directions while evading obstacles. By focusing on kinematics and obstacle grid representation, the method efficiently utilizes the versatility of a 7-degree-of-freedom (DOF) robot arm, employing positional and differential kinematics for movement calculations. The strategy doesn't require exhaustive training as with neural networks, making it deployable in various environments.

@@

Robotic manipulators with a high degree of freedom (DOF), possess a crucial characteristic known as kinematic redundancy~\cite{Nguyen1991}. Kinematic redundancy refers to having more joints (i.e., degrees of freedom - DOF) than required to perform a task, thus allowing the robot to adopt multiple configurations to accomplish the same end-effector position and orientation. This redundancy provides a considerable advantage, making it possible for manipulators to optimize movement according to different criteria, such as energy efficiency, obstacle avoidance, or joint limit avoidance.

@@

When dealing with redundant manipulators, task prioritization is crucial. In our approach, we distinguish between primary and secondary tasks. The primary task of the manipulator is usually defined as reaching a desired EE position and orientation or as following a path made out of such points. In this work we instead use a variation of APF (Artificial Potential Field), to guide the EE towards the goal position, while avoiding the obstacles in the path of the EE. Operating in a real-world environment often requires considering secondary tasks, such as real-time obstacle avoidance, to ensure safe and uninterrupted operation of the manipulator.

This paper proposes a novel method for real-time obstacle avoidance of the manipulator arm. We use an obstacle grid combined with convolution techniques to assign avoidance velocities to points of interest (POI) on manipulator segments, ensuring efficient and safe maneuvering within complex environments.

@@

The execution of the secondary task of obstacle avoidance necessitates an understanding of the spatial relationship between the manipulator and surrounding obstacles, usually quantified as a distance or interpreted as an avoidance or repulsive force. By conceptualizing obstacles as sources of repulsive forces, the manipulator can be guided away from potential collisions, akin to how two magnets of the same polarity repel each other.

Two commonly employed methods for obstacle avoidance include artificial potential field methods (APF)~\cite{Khatib1985} and artificial neural networks. In potential field methods, the robot and obstacles are treated as charged entities. The robot is attracted to the target (goal) while being repelled by obstacles, resulting in a potential field. The robot navigates this field, moving along the path of steepest descent. While this method is intuitive and relatively easy to implement, it can sometimes result in local minima problems where the robot gets stuck in a position that is not the target but cannot find a path due to the repelling obstacles. Another problem of the classical APF method is that it is focused on simple objects shapes and as such it is complicated to calculate repulsive field for more complex shapes of obstacles.

@@

The manipulation capabilities of a robot are fundamentally determined by its kinematics, the mathematical description of motion without considering the forces that cause it. In our experiments, we employed a 7 degree-of-freedom (DOF) robot. The seven joints provide a high level of versatility, enabling the robot to move and orient its end-effector in a broad range of positions and directions. To calculate the movements and positions of the robot, we use both positional and differential kinematics.

Positional kinematics focuses on the relationship between joint angles and the position and orientation of the end-effector. Forward kinematics involves calculating the position and orientation of the end-effector given the joint angles, while inverse kinematics is the process of determining the joint angles for a desired end-effector position and orientation. In the case or redundant manipulator, there isn't an unique mapping from EE coordinates to the joint coordinates as there exist an infinite number of different joint angles configurations that map to the same EE position.

@@

The null space (eq.~\ref{eq:Null-Space}) of the Jacobian corresponds to those joint motions which produce no movement in the end-effector. The primary task for our robot is the desired end-effector movement. The secondary task of obstacle avoidance in our research takes advantage of this null-space motion~\cite{lajpah_Petri_2012}. Our method applies an avoidance velocity from detected obstacles in the environment. We calculate this velocities at certain points of interest (POI) that are positioned on the manipulator arm segments. We transform avoidance velocities into the joint space velocities using $J^{\dagger}_{poi}$ Jacobian calculated for the specific POI. These velocities we than transform into the null-space using matrix $N$. The robot arm is thus able to carry out the primary task while simultaneously avoiding obstacles by appropriately utilizing its redundant DOF (eq.~\ref{eq:manipulator_kinematics}).

@@

We integrated a supplementary secondary task aiming to maintain the robot manipulator's joints as close to the midpoint of their range (centered between $q(i)_{max}$ and $q(i)_{min}$ for the i-th joint ) as possible ( eq.~\ref{eq:mid-joints}). $w(i)_m$ weights the importance of each joints position. This auxiliary task, designed to elevate the robot's "elbow", reinforces the robot's upright posture during the execution of the primary task and the obstacle avoidance task.

@@

@@

In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with kinematics. This methodology allows us to calculate the necessary avoidance velocities for specific points of interest located on the manipulator.

The process commences with positional kinematic transformations, which facilitate the mapping of points of interest from the robot's configuration space to the Cartesian space, followed by their projection onto the 3D voxel grid.

These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.

@@

Figure~\ref{y axis avoidance values} illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis. Here, the y-coordinates are presented on the x-axis, and the corresponding velocities are plotted on the y-axis. As expected, the y-velocity is the only component exhibiting non-zero values. It effectively propels the robot away from the wall, moving in either a positive or negative direction, depending on the relative position to the wall. Conversely, the x and z velocities remain zero, reflecting the absence of obstacles along these axes. This aligns with our expectations given the changing in presence of the wall obstacle is only along the y direction.

@@

The figure also highlights how the application of the avoidance algorithm prevents the fourth joint from moving closer to the wall, as compared to when the avoidance algorithm is not employed. The position of the joint, upon task completion, is noted to be approximately 20 cm further away from the wall. This distance depends on the specific choice of constants and kernel values.

@@

@@

Despite its effectiveness, the approach could potentially fall into local minima traps. This limitation could be addressed by incorporating multiple convolutional kernels, though this would increase complexity and resemble a structure of a convolutional neural networks.
